DNAm Hierarchical Model - Prior Sensitivity Analysis 

### 2020-10-20 prior_sensitivity.Rmd
updated prior analysis for figures

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
#library(gtools)
library(emdist)

@@

@@

    # BCD
    BC_temp[j] = -log(sum(sqrt(P$density*priordens)))

@@

print("N_eff")
kable(n_eff_mat, digits=2)
print("Rhat")
kable(Rhat_mat, digits=3)
print("KLdiv")
kable(KLdiv_mat, digits=3)
cat("\newpage")
print("BCdist")
kable(BCdist_mat, digits=3)
print("EMdist")
kable(EMdist_mat, digits=3)

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
#library(gtools)
library(emdist)
library(ggpubr)

@@

```{r prior_dataAndFits, echo=FALSE, comment=NA}
# load TCGA-based 450K array estimates (see estimateTCGApriors.R for data+fits)
load("/Users/kevinmurgas/Documents/Data+ project/TCGA data/estimates_allpats.Rdata")

# remove all zeros, and take square root of variance (stored variable) to get sigma/stdev
sigmaP_ests <- sqrt(sigmaP_ests[which(!!sigmaP_ests)])
sigmaPT_ests <- sqrt(sigmaPT_ests[which(!!sigmaPT_ests)])
sigmaT_ests <- sqrt(sigmaT_ests[which(!!sigmaT_ests)])
mu_ests <- mu_ests[which(!!mu_ests)]
betaT_ests <- betaT_ests[which(!!betaT_ests)]

gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
cols = gg_color_hue(4)

# ggplot histograms, with fit overlays
# mu
# 1x: 0.35570*normal(-3.0675,0.75214) + 0.64430*normal(1.1290, 1.3920)
# 2x: 0.35570*normal(-3.0675, 1.0637) + 0.64430*normal(1.1290, 1.9685)
# 3x: 0.35570*normal(-3.0675, 1.3027) + 0.64430*normal(1.1290, 2.4110)
# 5x: 0.35570*normal(-3.0675, 1.6818) + 0.64430*normal(1.1290, 3.1125)
f <- function(x, a, m1, s1, m2, s2) a*dnorm(x,m1,s1) + (1-a)*dnorm(x,m2,s2)
p1 <- ggplot(data.frame(p=mu_ests), aes(x=p)) + geom_histogram(binwidth=0.1,aes(y=..density..),fill="lightgray") + theme_classic() +
  #labs(x=expression(mu)) +
  xlim(-6,6) +
  stat_function(fun=function(x) f(x,0.3557,-3.0675, 0.75214,1.1290, 1.39197),geom="line",color=cols[1]) +
  stat_function(fun=function(x) f(x,0.3557,-3.0675, 1.0637,1.1290, 1.9685),geom="line",color=cols[2]) +
  stat_function(fun=function(x) f(x,0.3557,-3.0675, 1.3027,1.1290, 2.4110),geom="line",color=cols[3]) +
  stat_function(fun=function(x) f(x,0.3557,-3.0675, 1.6818,1.1290, 3.1125),geom="line",color=cols[4])
# betaT
# 1x: Cauchy(-0.04051, 0.19330)
# 2x: Cauchy(-0.04051, 0.38660)
# 3x: Cauchy(-0.04051, 0.57990)
# 5x: Cauchy(-0.04051, 0.96650)
p2 <- ggplot(data.frame(p=betaT_ests), aes(x=p)) + geom_histogram(binwidth=0.1,aes(y=..density..),fill="lightgray") + theme_classic() +
  #labs(x=expression(nu)) +
  xlim(-5,5) +
  stat_function(fun=function(x) dcauchy(x,-0.04051, 0.19330),geom="line",color=cols[1]) +
  stat_function(fun=function(x) dcauchy(x,-0.04051, 0.38660),geom="line",color=cols[2]) +
  stat_function(fun=function(x) dcauchy(x,-0.04051, 0.57990),geom="line",color=cols[3]) +
  stat_function(fun=function(x) dcauchy(x,-0.04051, 0.96650),geom="line",color=cols[4])
# sigmaE: all gamma(2,2)
p3 <- ggplot() + stat_function(fun=function(x) dgamma(x,2,2),geom="line",color="black") + xlim(0,2) + theme_classic() + labs(x=expression(sigma[E]))
# sigmaP
# 1x: gamma(5.1989, 16.6192)
# 2x: gamma(3.4016, 9.5057)
# 3x: gamma(2.7693, 7.0029)
# 5x: gamma(2.2298, 4.8674)
p4 <- ggplot(data.frame(p=sigmaP_ests), aes(x=p)) + geom_histogram(binwidth=0.02,aes(y=..density..),fill="lightgray") + theme_classic() +
  #labs(x=expression(sigma[P])) +
  xlim(0,2) +
  stat_function(fun=function(x) dgamma(x,5.1989, 16.6192),geom="line",color=cols[1]) +
  stat_function(fun=function(x) dgamma(x,3.4016, 9.5057),geom="line",color=cols[2]) +
  stat_function(fun=function(x) dgamma(x,2.7693, 7.0029),geom="line",color=cols[3]) +
  stat_function(fun=function(x) dgamma(x,2.2298, 4.8674),geom="line",color=cols[4])
# sigmaPT
# 1x: gamma(3.5539, 5.5324)
# 2x: gamma(2.5210, 3.2948)
# 3x: gamma(2.1457, 2.4819)
# 5x: gamma(1.8166, 1.7689)
p5 <- ggplot(data.frame(p=sigmaPT_ests), aes(x=p)) + geom_histogram(binwidth=0.02,aes(y=..density..),fill="lightgray") + theme_classic() +
  #labs(x=expression(sigma[PT])) +
  xlim(0,2) +
  stat_function(fun=function(x) dgamma(x,3.5539, 5.5324),geom="line",color=cols[1]) +
  stat_function(fun=function(x) dgamma(x,2.5210, 3.2948),geom="line",color=cols[2]) +
  stat_function(fun=function(x) dgamma(x,2.1457, 2.4819),geom="line",color=cols[3]) +
  stat_function(fun=function(x) dgamma(x,1.8166, 1.7689),geom="line",color=cols[4])
# sigmaT
# 1x: gamma(6.8383, 25.7560)
# 2x: gamma(4.2574, 14.370)
# 3x: gamma(3.3643, 10.430)
# 5x: gamma(2.6144, 7.1221)
p6 <- ggplot(data.frame(p=sigmaT_ests), aes(x=p)) + geom_histogram(binwidth=0.02,aes(y=..density..),fill="lightgray") + theme_classic() +
  xlim(0,2) +
  #labs(x=expression(sigma[T])) +
  stat_function(fun=function(x) dgamma(x,6.8383, 25.7560),geom="line",color=cols[1]) +
  stat_function(fun=function(x) dgamma(x,4.2574, 14.370),geom="line",color=cols[2]) +
  stat_function(fun=function(x) dgamma(x,3.3643, 10.430),geom="line",color=cols[3]) +
  stat_function(fun=function(x) dgamma(x,2.6144, 7.1221),geom="line",color=cols[4])

p_all <- ggarrange(p1,p2,p3,p4,p5,p6)
# ggsave("prior_fits_all.svg",p_all,height=3,width=5.53,dpi=300)
```

@@

    # BCD
    dx <- x[2]-x[1]
    BC_temp[j] = -log(sum(sqrt(P$density*priordens)*dx))

@@

print("N_eff")
kable(n_eff_mat, digits=2)
print("Rhat")
kable(Rhat_mat, digits=3)
print("KLdiv")
kable(KLdiv_mat, digits=3)
cat("\n")
print("BCdist")
kable(BCdist_mat, digits=3)
print("EMdist")
kable(EMdist_mat, digits=3)

@@

\newpage
Plots
```{r plots, echo=FALSE, comment=NA}
# goal here is to just get the distributions of parameters for each prior set
# store histogram density, and KDE
# plot histogram with side-by-side bar plots
# plot KDE's as line plots

mu_allprior <- data.frame(x1=numeric(866091),
                          x2=numeric(866091),
                          x3=numeric(866091),
                          x5=numeric(866091))
betaT_allprior <- mu_allprior
sigmaP_allprior <- mu_allprior
sigmaPT_allprior <- mu_allprior
sigmaT_allprior <- mu_allprior
sigmaE_allprior <- mu_allprior

  mu_allprior[,i] <- mu_full$p50
  betaT_allprior[,i] <- betaT_full$p50
  sigmaP_allprior[,i] <- sigmaP_full$p50
  sigmaPT_allprior[,i] <- sigmaPT_full$p50
  sigmaT_allprior[,i] <- sigmaT_full$p50
  sigmaE_allprior[,i] <- sigmaE_full$p50
}

cat("Posterior estimates for varying prior relaxation")
pM <- mu_allprior %>% gather("relax","val",1:4) %>% ggplot(aes(x=val,y=..density..,color=relax)) + geom_density() + xlim(-6,6) + ggtitle("Mu") + theme_classic() + theme(legend.position = "none")
pB <- betaT_allprior %>% gather("relax","val",1:4) %>% ggplot(aes(x=val,y=..density..,color=relax)) + geom_density() + xlim(-5,5) + ggtitle("BetaT") + theme_classic() + theme(legend.position = "none")
pP <- sigmaP_allprior %>% gather("relax","val",1:4) %>% ggplot(aes(x=val,y=..density..,color=relax)) + geom_density() + xlim(0,2) + ggtitle("sigmaP") + theme_classic() + theme(legend.position = "none")
pPT <- sigmaPT_allprior %>% gather("relax","val",1:4) %>% ggplot(aes(x=val,y=..density..,color=relax)) + geom_density() + xlim(0,2) + ggtitle("sigmaPT") + theme_classic() + theme(legend.position = "none")
pT <- sigmaT_allprior %>% gather("relax","val",1:4) %>% ggplot(aes(x=val,y=..density..,color=relax)) + geom_density() + xlim(0,2) + ggtitle("sigmaT") + theme_classic() + theme(legend.position = "none")
pE <- sigmaE_allprior %>% gather("relax","val",1:4) %>% ggplot(aes(x=val,y=..density..,color=relax)) + geom_density() + xlim(0,2) + ggtitle("sigmaE") + theme_classic() + theme(legend.position = "none")
pLP <- sigmaP_allprior %>% gather("relax","val",1:4) %>% mutate(val=log(val)) %>% ggplot(aes(x=val,y=..density..,color=relax)) + geom_density() + xlim(-4,1) + ggtitle("log(sigmaP)") + theme_classic() + theme(legend.position = "none")
pLPT <- sigmaPT_allprior %>% gather("relax","val",1:4) %>% mutate(val=log(val)) %>% ggplot(aes(x=val,y=..density..,color=relax)) + geom_density() + xlim(-4,1) + ggtitle("log(sigmaPT)") + theme_classic() + theme(legend.position = "none")
pLT <- sigmaT_allprior %>% gather("relax","val",1:4) %>% mutate(val=log(val)) %>% ggplot(aes(x=val,y=..density..,color=relax)) + geom_density() + xlim(-4,1) + ggtitle("log(sigmaT)") + theme_classic() + theme(legend.position = "none")

p_all<- ggarrange(pM, pB, pE, pP, pPT, pT, pLP, pLPT, pLT, nrow=3, ncol=3)
print(p_all)
# ggsave("prior_sens_kdes.svg",p1)
```
